---
title: "Run_Crop2ML_DSSAT_SoilTemp"
author: "RCichota"
date: "`r Sys.Date()`"
output: html_document
---


# Running a soil temperature model from Crop2ML in R - DSSAT

This is part of a series to demonstrate how to run in R environment a soil temperature model generated using Crop2ML.

Crop2ML (Crop Modelling Meta Language) is a framework to translate components from and to specific modeling platforms. It provides a solution that can generate components automatically, based on a metalanguage, to simulate particular soil and/or plant processes. The framework was proposed by the Agricultural Model Exchange Initiative (AMEI) to facilitate the exchange and reuse of model components between different process-based modeling platforms used in agricultural and environmental research. More info at: https://crop2ml.org

A version of the native soil temperature model from the DSSAT platform (https://dssat.net) was converted into R functions using Crop2ML (working version as of Oct/2024). The scripts below show how to use the model stand-alone in R. This is a simple demonstration on how to set up and run the model for a very simple example, as well as record and graph some results from this example. More elaborate exercises and coupling with other models are possible but not in scope here.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load some libraries
library(here)
library(gsubfn)
library(dplyr)
library(lubridate)
library(ggplot2)

# load the DSSAT native component
source(here("Crop2ML", "DSSAT_ST_standalone", "STEMP_Component.r"))

```

## Read in parameters and basic data

The soil temperature requires a number of soil parameters and input data (weather). These are read in from test files.
Note that some model specific parameters are set in here manually...

```{r read files}

# read general site info (weather and soil type summary)
siteFilePath <- here("data", "SiteExample.dat")
siteData <- read.table(siteFilePath, skip = 2, header = TRUE)

# get the weather data (daily values)
weatherFilePath <- here("data", "WeatherDataExample.dat")
weatherData <- read.table(weatherFilePath, skip = 2, header = TRUE)

# get soil profile data
soilLayerFilePath <- here("data", "SoilLayersExample.dat")
soilLayerData <- read.table(soilLayerFilePath, skip = 2, header = TRUE)

```

## Run the model over a number of dates

Firstly the dates to start and end the period over which the simulation will run are defined, then the soil water content and some model specific variables/parameters. Finally the model is run (loop over the weather data)
Here we are collecting a few variables of interest, putting in a data.frame with names and formats wanted (assumed), and then saving it to a file in the outputs folder.
The example simulation runs over one year, the data collected consists of soil temperature for the various depths (as defined in soilLayerData) for each simulated date.

```{r run simulation}

# set the dates between which the model will run over
startDate <- "2000-01-01"
endDate <- "2000-12-31"
firstDay <- match(startDate, weatherData$date)
lastDay <- match(endDate, weatherData$date)

# define some variables specific to a given run/simulation
fractionPAW <- 0.75
currentSWC <- soilLayerData$LLC + (soilLayerData$DUL - soilLayerData$LLC)*fractionPAW

doWater <- "Y"     # soil water switch

# create some variables to collect outputs of interest
simulatedDate <- c()
simulatedDepth <- c()
simulatedTemperature <- c()

# set model initialisation
initialState <- init_stemp(
  NL=nrow(soilLayerData),
  ISWWAT=doWater,
  BD=soilLayerData$BD,
  DLAYR=soilLayerData$Thickness,
  DS=soilLayerData$LayerBot,
  DUL=soilLayerData$DUL,
  LL=soilLayerData$LLC,
  NLAYR=nrow(soilLayerData),
  MSALB=siteData$SAlbedo,
  SRAD=weatherData$radn[firstDay],
  SW=currentSWC,
  TAVG=(weatherData$mint[firstDay]+weatherData$maxt[firstDay])/2,
  TMAX=weatherData$maxt[firstDay],
  XLAT=siteData$Latitude,
  TAV=siteData$TAV,
  TAMP=siteData$TAMP,
  DOY=yday(as.Date(weatherData$date[firstDay])))

# collect some variables that are needed for next iteration
previousCUMDPT <- initialState$CUMDPT
previousDSMID <- initialState$DSMID
previousTMA <- initialState$TMA
previousTDL <- initialState$TDL
previousATOT <- initialState$ATOT
previousSRFTEMP <- initialState$SRFTEMP
previousST <- initialState$ST

# run model between two days (dates in weather file)
for (day in firstDay:lastDay)
{
  onProcessState <- model_stemp_(
    TMAX=weatherData$maxt[day],
    MSALB=siteData$SAlbedo,
    HDAY=initialState$HDAY,
    NL=nrow(soilLayerData),
    SRFTEMP=previousSRFTEMP,
    LL=soilLayerData$LLC,
    NLAYR=nrow(soilLayerData),
    DS=soilLayerData$LayerBot,
    ST=previousST,
    SRAD=weatherData$radn[day],
    TAMP=siteData$TAMP,
    DLAYR =soilLayerData$Thickness,
    TMA=previousTMA,
    TDL=previousTDL,
    ISWWAT=doWater,
    CUMDPT=previousCUMDPT,
    TAVG=(weatherData$mint[day]+weatherData$maxt[day])/2,
    BD=soilLayerData$BD,
    SW=currentSWC,
    ATOT=previousATOT,
    TAV=siteData$TAV,
    XLAT=siteData$Latitude,
    DUL=soilLayerData$DUL,
    DSMID=previousDSMID,
    DOY=yday(as.Date(weatherData$date[day])))

  # collect some variables that are needed for next iteration
  previousCUMDPT <- onProcessState$CUMDPT
  previousDSMID <- onProcessState$DSMID
  previousTMA <- onProcessState$TMA
  previousTDL <- onProcessState$TDL
  previousATOT <- onProcessState$ATOT
  previousSRFTEMP <- onProcessState$SRFTEMP
  previousST <- onProcessState$ST

  # collect output data
  simulatedDate <- c(simulatedDate, rep(weatherData$date[day], nrow(soilLayerData) + 1))
  simulatedDepth <- c(simulatedDepth, c(0.0, (soilLayerData$LayerTop+soilLayerData$LayerBot)/2))
  simulatedTemperature <- c(simulatedTemperature, c(round(onProcessState$SRFTEMP, 6), round(onProcessState$ST, 6)))
}

# prepare the output table
outputData <- data.frame(simulatedDate, simulatedDepth, simulatedTemperature)
colnames(outputData) <- c("Date", "Depth", "Temperature")

# save the outputs
thisModel <- "DSSAT"
outputFileName <- paste0("OutputDataExample_", thisModel, ".csv")
outputFilePath <- here("outputs", outputFileName)
write.table(outputData, outputFilePath, sep=",", row.names=FALSE, quote=FALSE)

```

## Graph some results

Using the data.frame created above (otherwise one has to read in the file saved), two graphs are created to check the model and demonstrate how the data can be presented. The first graph shows the time series of simulated temperature at three depths; the second graph shows the temperature simulated over the profile (all depth) for four dates along the year.

```{r graphs}

timeSeriesTemperature <- outputData %>%
  filter(Depth==0 | Depth==100 | Depth==1125) %>%
  mutate_at(vars(Depth), factor)

ggplot(data=timeSeriesTemperature, aes(x=as.Date(Date), y=Temperature, color=Depth)) +
  geom_line() +
  labs(title = "Simulated temperature over time at different depths", y = "Temperature (oC)", x = "Date")


profileTemperature <- outputData %>%
  filter(Date == "2000-01-01" | Date == "2000-04-01" | Date == "2000-07-01" | Date == "2000-10-01")

ggplot(data=profileTemperature, aes(x=Temperature, y=Depth, color=Date)) +
  geom_path() +
  scale_y_reverse() +
  labs(title = "Simulated temperature profile at different dates", y = "Depth (mm)", x = "Temperature (oC)")

```
